Wellcome Open Research Article Template

### 2023-02-10 manuscript/manuscript.tex
update results in paper

Every week, we submitted an ensemble of individual forecasts ("crowd forecast") to the European Forecast Hub. In contrast to the ensemble of human forecasts described in \citet{bosseComparingHumanModelbased2022}, we used the quantile-wise median, rather than the quantile-wise mean to combine predictions. %, following suggestions made in CITATION. 
We submitted three different ensembles: The first one, "epiforecasts-EpiExpert\_direct" (here called "Crowd direct") was a quantile-wise median ensemble of all the direct forecasts. "epiforecasts-EpiExpert\_Rt" (here called "Crowd Rt" was a median ensemble of all forecasts made through the $R_t$ interface. "epiforecasts-EpiExpert" (here called "Crowd combined") was a median ensemble of all forecasts, after combining direct and $R_t$ forecast from the same individual by taking a quantile-wise mean. Before creating the ensemble, we deleted forecasts that were clearly the result of a user or software error (such as forecasts that were zero everywhere).

@@

Our analysis of the data largely follows the one outlined in \citet{bosseComparingHumanModelbased2022}. If not otherwise stated, we present results for two-week-ahead forecasts, following the practice adopted by the COVID-19 Forecast Hubs that found poor predictive performance beyond this horizon [15].

We scored forecasts using the weighted interval score (WIS, formulas given in the SI in Section \ref{sec:wis}) \cite{bracherEvaluatingEpidemicForecasts2021} and used the empirical coverage of all central 50\% and 90\% prediction intervals to measure probabilistic callibration. The WIS is proper scoring rule (with smaller values implying better performance). A forecaster, in expectation, optimises her score by providing a predictive distribution $F$ that is equal to the data-generating distribution $G$ and is therefore incentivised to report her true belief. The WIS can be seen as a generalisation of the absolute error to predictive distribution in a quantile-based format and represents an approximation of the continuous ranked probability score (CRPS). The WIS can be decomposed into three separate penalty components: forecast spread (i.e. uncertainty of forecasts), over-prediction and under-prediction. FORMULAS IN APPENDIX. Empirical coverage means the percentage of observations falling inside any given prediction interval (e.g. the cumulative percentage of observed values that fall inside all 50\% prediction intervals). It represents a measure of probabilistic calibration \citep{gneitingProbabilisticForecastsCalibration2007}. Forecasts were evaluated using the \texttt{scoringutils} \citep{bosseEvaluatingForecastsScoringutils2022} package in \textsf{R}.

@@

For two-week-ahead-forecasts of incident cases, the Forecast Hub ensemble outperformed all crowd forecasting ensembles as measured by the weighted interval score (WIS). On the natural scale, the WIS as a measure of the absolute distance between forecast and observation increases or decreases with the magnitude of the forecast target \cite{bosseTransformationForecastsEvaluating2023, bracherEvaluatingEpidemicForecasts2021}. Average scores are therefore dominated by performance around the peak when cases were highest, in particular by forecasts made on the 19th of July for the 31st of July \ref{fig:forecasts-scores}. Direct forecasts tended to be higher throughout the study period and especially during the peak than both the $R_t$ forecasts and the Forecast Hub ensemble and therefore showed the poorest performance (see Figure \ref{fig:performance} and Table \ref{tab:scores}). When looking at WIS values on the log scale (that is, when log-transforming forecasts and observations prior to applying the WIS), scores are more equally distributed across the study period and more weight is given to forecasts in June and July which underpredicted the extent to which case number would rise. On the log scale, the Forecast Hub ensemble performs slightly worse than all of the human forecasting approaches. Similarly to what was observed previously \citep{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a}, all crowd forecasts as well the Forecast Hub ensemble, exhibited underdispersion, meaning that forecasts on average were too narrow and not uncertain enough, with empirical coverage below nominal coverage.

\begin{table}[!h]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{llcccc}
\toprule
Model & Target & WIS & WIS (log scale) & Coverage 50\% & Coverage 90\%\\
\midrule
EuroCOVIDhub-ensemble & Cases & 38000 & 0.25 & 0.38 & 0.69\\
crowd-direct & Cases & 46000 & 0.23 & 0.46 & 0.62\\
crowd-ensemble & Cases & 40000 & 0.22 & 0.38 & 0.69\\
crowd-rt & Cases & 39000 & 0.24 & 0.31 & 0.69\\
\addlinespace
EuroCOVIDhub-ensemble & Deaths & 57 & 0.16 & 0.62 & 0.85\\
crowd-direct & Deaths & 78 & 0.17 & 0.46 & 0.77\\
crowd-ensemble & Deaths & 61 & 0.16 & 0.38 & 0.77\\
crowd-rt & Deaths & 63 & 0.23 & 0.23 & 0.69\\
\bottomrule
\end{tabular}}
\caption{Table caption}
\label{tab:scores}
\end{table}

@@

In terms of the WIS the Forecast Hub ensemble also outperformed all human forecasts when forecasting deaths. This is true as well for WIS values obtained after applying a log-transformation. From June to the end of July, direct forecasts of deaths were noticeable higher than $R_t$ death forecasts, whereas in August direct death forecasts were substantially lower than $R_t$ death forecasts. On the natural scale, the largest scores occured at the end of the study period when deaths were highest. Direct forecasts, which most strongly underpredicted deaths in August therefore showed the poorest performance overall. On the log scale, however, underprediction in June and July when death incidences were lower, received more weight and direct crowd forecast outperformed $R_t$ forecasts. \textbf{CHECK WITH OTHER GROUND TRUTH DATA?}. Empirical coverage for deaths seems broadly comparable for all human forecasts, whereas the Forecast Hub ensemble displays noticeably better coverage for deaths than for cases (Figure \ref{fig:performance}E, F).

@@

@@

- hard to draw conclusions, performance changes depending on how you look at the data. 
- headline result not supported? 
- still some evidence that forecasting deaths is easier for computers
-

>>>>

Every week, we submitted an ensemble of individual forecasts ("crowd forecast") to the European Forecast Hub. In contrast to the ensemble of human forecasts described in \citet{bosseComparingHumanModelbased2022}, we used the quantile-wise median, rather than the quantile-wise mean to combine predictions. %, following suggestions made in CITATION. 
We submitted three different ensembles: The first one, "epiforecasts-EpiExpert\_direct" (here called "Crowd direct") was a quantile-wise median ensemble of all the direct forecasts. "epiforecasts-EpiExpert\_Rt" (here called "Crowd Rt" was a median ensemble of all forecasts made through the $R_t$ interface. "epiforecasts-EpiExpert" (here called "Crowd combined") was a median ensemble of all forecasts together, meaning that a participant could enter the the median ensemble twice if she had submitted both a direct and an $R_t$ forecast. Before creating the ensemble, we deleted forecasts that were clearly the result of a user or software error (such as forecasts that were zero everywhere).

@@

Our analysis of the data largely follows the one outlined in \citet{bosseComparingHumanModelbased2022}. If not otherwise stated, we present results for two-week-ahead forecasts, following the practice adopted by the COVID-19 Forecast Hubs that found poor predictive performance beyond this horizon [15]. We analysed all forecasts together separated by forecasting forecast target (cases or deaths). In addition, we compared the performance of the direct vs. $R_t$ forecasting approach using instances where we had both a direct forecasts and an $R_t$ forecast from the same person. Those forecasts were combined using a quantile-wise median separately for direct and $R_t$ forecasts such that we obtained a modified direct human ensemble and a modified $R_t$ ensemble that included corresponding forecasts from the same forecaster.

We scored forecasts using the weighted interval score (WIS, formulas given in the SI in Section \ref{sec:wis}) \cite{bracherEvaluatingEpidemicForecasts2021} and used the empirical coverage of all central 50\% and 90\% prediction intervals to measure probabilistic callibration. The WIS is proper scoring rule (with smaller values implying better performance). A forecaster, in expectation, optimises her score by providing a predictive distribution $F$ that is equal to the data-generating distribution $G$ and is therefore incentivised to report her true belief. The WIS can be seen as a generalisation of the absolute error to predictive distribution in a quantile-based format and represents an approximation of the continuous ranked probability score (CRPS). The WIS can be decomposed into three separate penalty components: forecast spread (i.e. uncertainty of forecasts), over-prediction and under-prediction. Empirical coverage means the percentage of observations falling inside any given prediction interval (e.g. the cumulative percentage of observed values that fall inside all 50\% prediction intervals). It represents a measure of probabilistic calibration \citep{gneitingProbabilisticForecastsCalibration2007}. Forecasts were evaluated using the \texttt{scoringutils} \citep{bosseEvaluatingForecastsScoringutils2022} package in \textsf{R}.

@@

For two-week-ahead-forecasts of incident cases, the Forecast Hub ensemble outperformed all crowd forecasting ensembles as measured by the weighted interval score (WIS). On the natural scale, the WIS as a measure of the absolute distance between forecast and observation increases or decreases with the magnitude of the forecast target \cite{bosseTransformationForecastsEvaluating2023, bracherEvaluatingEpidemicForecasts2021}. Average scores are therefore dominated by performance around the peak when cases were highest, in particular by forecasts made on the 19th of July for the 31st of July \ref{fig:forecasts-scores}.$R_t$ forecasts tended to be higher than both the direct forecasts and the Forecast Hub ensemble, especially around the peak, and therefore showed the poorest performance in terms of WIS on the natural scale (see Figure \ref{fig:performance} and Table \ref{tab:scores}). When looking at WIS values on the log scale,
%(that is, when log-transforming forecasts and observations prior to applying the WIS), 
scores are more equally distributed across the study period and more weight is given to forecasts in June and July which underpredicted the extent to which case number would rise. On the log scale, all human forecast ensembles performed sligthly better than the Forecast Hub ensemble. Similarly to what was observed previously for case forecasts \citep{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a}, all forecasting approaches exhibited underdispersion, meaning that forecasts on average were too narrow and not uncertain enough, with empirical coverage below nominal coverage. This was the case most severely for the $R_t$ forecasts.

\input performance-table.tex

@@

In terms of the WIS on the natural scale, direct human forecasts performed best, while $R_t$ forecasts performed significantly worse than all other approaches (see Figure \ref{fig:performance}B and Table \ref{tab:scores}). From June to the end of July, $R_t$ forecasts overpredicted deaths and were noticeable higher than other forecasts, whereas in August $R_t$ forecasts underpredicted deaths and were substantially lower than other forecasts. On the natural scale, the highest scores were concentrated towards the end of the study period, where death incidences were highest, while on the log scale, scores were more evenly distributed across the study period. On the log scale, direct human forecasts, the combined crowd ensemble and the Forecast Hub ensemble showed comparable performance, while $R_t$ forecasts performed again significantly worse. Note that the picture changes when looking at the revised data which was corrected upwards compared to the original data(DO WE WANT TO DO THAT? IS THAT MEANINGFUL?) (see Figure \ref{fig:data-revisions}). For the revised data, direct forecasts performed significantly worse than both the Forecast Hub ensemble and the $R_t$ forecasts on the log scale, while still performing better than the $R_t$ forecasts on the natural scale (see Table \ref{tab:scores-revised}). 
%For revised data: On the natural scale, the largest scores occured at the end of the study period when deaths were highest. On the log scale, however, underprediction in June and July when death incidences were lower, received more weight and direct crowd forecast outperformed $R_t$ forecasts. 
$R_t$ forecasts show strong underdispersion with empirical coverage of the 50\% prediction intervals being close to zero and smaller than 50\% for the 90\% prediction intervals (see Figure \ref{fig:performance} and Table \ref{tab:scores}). The Forecast Hub ensemble was overdispersed with empirical coverage exceeding nominal coverage, and for the direct human forecasts, empirical coverage was close to nominal coverage.

% For data revision: deaths seems broadly comparable for all human forecasts, whereas the Forecast Hub ensemble displays noticeably better coverage for deaths than for cases (Figure \ref{fig:performance}E, F).

@@

\subsection*{Rt forecasts}

\begin{figure}[H]
\centering
\includegraphics[width=0.99\textwidth]{../output/figures/comparison-direct-rt-individual.png}
\caption{\bf{Change in WIS for forecasters making either direct or $R_t$ forecasts. The absolute level for a given forecaster compared to others is not meaningful as forecasters differ in the amounts of forecasts they have submitted.}}
\label{fig:comparison-direct-rt-individual}
\end{figure}

For cases, where participants could observe the case forecast implied their $R_t$ forecast, predictive performance was broadly comparable between direct and $R_t$ forecasts for most forecasters (see Figure \ref{comparison-direct-rt-individual}). For deaths, predictive performance was significantly worse for most forecasters, suggesting that the convolution from latent infections implied by the $R_t$ forecasts to deaths did not work well. IMPORTANTLY THOUGH, THE PICTURE DOES LOOK DIFFERNT WHEN LOOKING AT THE REVISED DATA. When looking at an ensemble including only those data points where a single forecaster has submitted both a direct and an $R_t$ forecast, we found a similar picture. Performance for cases was broadly comparable, with direct forecasts slighly better than $R_t$ forecasts on the natural scale and slightly worse on the log scale. On the log scale, the ensemble of $R_t$ forecasts performed significantly worse than the ensemble of corresponding direct forecasts both on the natural and on the log scale.

@@

- hard to draw conclusions, performance changes depending on how you look at the data. Different horizons look different. 
- headline result not supported? 
- still some evidence that forecasting deaths is easier for computers
-
